## Loading and preprocessing the data

First load the data

```{r}
df_data <- read.csv("C:\\Users\\aozbay\\Desktop\\Reproducible Research\\activity.csv")
head(df_data)
```

Looks like there are NAs in the steps column. Remove the records with NAs for the time being. We will do a better job of imputing later.

```{r}
df_data <- df_data[complete.cases(df_data),]
head(df_data)
```


## What is mean total number of steps taken per day?

Let's first calculate the stepstaken each day

```{r}
steps.daily <- aggregate(df_data$steps, by=list(df_data$date), FUN=sum)
colnames(steps.daily) <- c("date","steps")
head(steps.daily)
```

Let's take a look at the historgram

```{r}
hist(steps.daily$steps)
```

What are the mean and median of steps?

```{r}
steps.summary <- summary(steps.daily$steps)
steps.median <- steps.summary[3]
steps.mean <- steps.summary[4]
print(steps.summary)
```

The median is `r steps.median` and the mean is `r steps.mean`.

## What is the average daily activity pattern?

First let's find the average number of steps taken each 5 minute period

```{r}
steps.five_minperiod <- aggregate(df_data$steps, list(df_data$interval), mean)
colnames(steps.five_minperiod ) <- c("interval", "avg.steps")
head(steps.five_minperiod)
```

Then let's plot it to see if there is a pattern

```{r}
plot(y=steps.five_minperiod$avg.steps, x=steps.five_minperiod$interval, type="l"
    ,xlab="Interval"
    ,ylab="Average Steps")
```

```{r echo=FALSE}
steps.interval_max_avg_steps <- 
  steps.five_minperiod[order(steps.five_minperiod$avg.steps, decreasing=TRUE)[1],]
```

The 5-minute interval, on average across all the days, that contains the maximum number of steps is `r steps.interval_max_avg_steps$interval` with `r steps.interval_max_avg_steps$avg.steps` average steps.

## Imputing missing values

We originally removed records with NAs. Let's do a better job by imputing them. First load the data again.

```{r}
df_data_new <- read.csv("C:\\Users\\aozbay\\Desktop\\Reproducible Research\\activity.csv")
head(df_data_new)
steps_new.num_missing_vals <- sum(is.na(df_data_new$steps))
```

There are `r steps_new.num_missing_vals` missing values. We already have the mean values for each interval. Let's substitute these means for the missing values.

```{r}
df_data_temp <- df_data_new[is.na(df_data_new$steps),]
df_data_temp <- merge(df_data_temp, steps.five_minperiod, by="interval")
df_data_temp$steps <- df_data_temp$avg.steps
df_data_temp$avg.steps <- NULL
df_data_new <- rbind(df_data_new[!is.na(df_data_new$steps),], df_data_temp)
df_data_new <- df_data_new[order(df_data_new$date, df_data_new$interval),]
rownames(df_data_new) <- 1:nrow(df_data_new)
head(df_data_new)
```

Now let's recalculate the stepstaken each day

```{r}
steps_new.daily <- aggregate(df_data_new$steps, by=list(df_data_new$date), FUN=sum)
colnames(steps_new.daily) <- c("date","steps")
head(steps_new.daily)
```

Let's take a look at the historgram

```{r}
hist(steps_new.daily$steps)
```

What are the mean and median of steps?

```{r}
steps_new.summary <- summary(steps_new.daily$steps)
steps_new.median <- steps_new.summary[3]
steps_new.mean <- steps_new.summary[4]
print(steps_new.summary)
```

The new median is `r steps_new.median` and the new mean is `r steps_new.mean`.
Previously the median was `r steps.median` and the mean was `r steps.mean`.

Only the median value changed after the imputation, as expected.


## Are there differences in activity patterns between weekdays and weekends?

In order to find any patterns between weekdays and weekends, let's modify our data frame a little.

```{r}
df_data_new$day_of_week <- weekdays(as.Date(df_data_new$date))
df_data_new$day_label <- "weekday"
df_data_new[(df_data_new$day_of_week == "Saturday" | df_data_new$day_of_week == "Sunday"),"day_label"] <- "weekend"
df_data_new$day_of_week <- NULL
head(df_data_new)
```

Let's calculate the average steps by interval by day_label

```{r}
avg_steps <- aggregate(df_data_new$steps, list(df_data_new$interval, df_data_new$day_label), FUN=mean)
colnames(avg_steps) <- c("interval", "day_label", "avg_steps")
par(mfcol=c(2,1))
plot(y=avg_steps[avg_steps$day_label=="weekday","avg_steps"],
     x=avg_steps[avg_steps$day_label=="weekday","interval"],
     type="l",
     xlab="Interval",
     ylab="Average Steps",
     sub="Weekday")
plot(y=avg_steps[avg_steps$day_label=="weekend","avg_steps"],
     x=avg_steps[avg_steps$day_label=="weekend","interval"],
     type="l",
     xlab="Interval",
     ylab="Average Steps",
     sub="Weekend")
```


