# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

First load the data

```{r}
df_data <- read.csv("C:\\Users\\aozbay\\Desktop\\Reproducible Research\\activity.csv")
head(df_data)
```

Looks like there are NAs in the steps column. Chnage them to 0s.

```{r}
df_data[is.na(df_data$steps),"steps"] <- 0
head(df_data)
```

## What is mean total number of steps taken per day?

Let's first calculate the stepstaken each day

```{r}
steps.daily <- aggregate(df_data$steps, by=list(df_data$date), FUN=sum)
colnames(steps.daily) <- c("date","steps")
head(steps.daily)
```

Let's take a look at the historgram

```{r}
hist(steps.daily$steps)
```

What are the mean and median of steps?

```{r}
steps.summary <- summary(steps.daily$steps)
steps.median <- steps.summary[3]
steps.mean <- steps.summary[4]
print(steps.summary)
```

The median is `r steps.median` and the mean is `r steps.mean`.

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
